### 2017-08-14 threshold/draft.tex
Updated draft following Mike's revision
Switching to the tex version only--no longer using Sweave (draft.tex is the working document now)

\documentclass{article}

\usepackage{Sweave}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{cite}
\usepackage{authblk}
\usepackage{graphicx}
\let\proglang=\textsf

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}

\newenvironment{proof}[1][Proof]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newenvironment{definition}[1][Definition]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newenvironment{example}[1][Example]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newenvironment{remark}[1][Remark]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}

\newcounter{algorithmctr}
\newenvironment{algorithm}{
   \refstepcounter{algorithmctr}
   \bigskip\noindent
   \textbf{Algorithm \thealgorithmctr\\}
}
{\bigskip}
\numberwithin{algorithmctr}{section}

\title{Efficient Thresholded Correlation using Truncated Singular Value Decomposition}
%\author{
%James Baglama\\
%University of Rhode Island
%\And
%Michael Kane\\
%Yale University
%\AND
%Bryan Lewis\\
%Paradigm4, Inc.\\
%\And
%Alex Poliakov\\
%Paradigm4, Inc.
%}
%\Plainauthor{James Baglama, Michael Kane, Bryan Lewis, Alex Poliakov}
%\Plaintitle{Efficient Thresholded Correlation using Truncated Singular Value Decomposition}
%\Shorttitle{Efficient Thresholded Correlation}
\author[1]{James Baglama \thanks{jbaglama@math.uri.edu}}
\author[2]{Michael Kane \thanks{michael.kane@yale.edu}}
\author[3]{Bryan Lewis \thanks{blewis@illposed.net}}
\author[3]{Alex Poliakov \thanks{apoliakov@paradigm4.com}}
\affil[1]{
University of Rhode Island
}
\affil[2]{
Yale University
}
\affil[3]{
Paradigm4, Inc.
}
\renewcommand\Authands{ and }

%\Abstract{
\begin{document}
\input{draft-concordance}
\maketitle
\begin{abstract}
Efficiently computing a subset of a correlation matrix consisting of values above 
a specified threshold is important to many practical applications of large-scale 
statistics and machine learning. Real-world problems in genomics, machine 
learning, finance, and others can produce correlation matrices too large to 
explicitly form and to tractably compute. However, in most applications, only 
values corresponding to highly-correlated (or anti-correlated) vectors are of 
interest, and those corresponding values typically make up a small fraction of 
the overall correlation matrix. For this situation, we present a method based on 
the singular value decomposition (SVD) and its relationship to the data 
covariance structure for efficiently computing thresholded subsets of very large correlation matrices.
\end{abstract}
%\Keywords{thresholded correlation, SVD, \proglang{R}}
%\Plainkeywords{thresholded correlation, SVD, R}
%\Address{
%James Baglama, Chair of Mathematics\\
%Department of Mathematics\\
%University of Rhode Island\\
%Kingston, Rhode Island 02881\\
%E-mail: \email{jbaglama@math.uri.edu}\\
%URL: \url{http://www.math.uri.edu/~jbaglama}\\
%\\
%Michael Kane\\
%Yale University\\
%300 Geroge Street\\
%New Haven, Connecticut 06511\\
%E-mail: \email{michael.kane@yale.edu}\\
%\\
%Bryan Lewis\\
%Paradigm4, Inc.\\
%281 Winter Street, Suite 360\\
%Waltham, MA 02451\\
%E-mail: \email{blewis@illposed.net}\\
%URL: \url{http://illposed.net}\\
%\\
%Alex Poliakov\\
%Paradigm4, Inc.\\
%281 Winter Street, Suite 360\\
%Waltham, MA 02451\\
%E-mail: \email{apoliakov@paradigm4.com}\\
%}

\section{Introduction}\label{intro}

Finding highly-correlated pairs among a large set of vectors is an important part 
of many statistics and machine learning applications. For instance, subsets of 
highly-correlated vectors of gene expression values may be used in the discovery 
of networks of genes relevant to particular biological processes \cite{genomics}. 
Identification of highly-correlated pairs may be used in feature selection 
algorithms for model fitting \cite{ml1, ml2} and are important to time series and 
image processing applications \cite{timeseries, svd-similarity}.

Because the number of correlation coefficients grows quadratically with the 
number of vectors, simply computing all pairs of correlation coefficients and 
then filtering out coefficients below a given threshold may not be 
computationally feasible for high-dimensional data. As a result, 
considerable attention has been devoted to \emph{pruning methods} that cheaply 
prune away all but the most highly-correlated pairs of vectors.  For example the 
TAPER and related algorithms of Xiong et al. develop pruning rules based on the 
frequency of mutually co-occurring observations (rows) among vector pairs; see 
\cite{prune1, prune2} and the references therein.  Related methods approximate a 
set containing the most highly-correlated pairs using hashing algorithms 
\cite{prune3}. Other methods arrive at approximate sets of the most 
highly-correlated pairs using dimensionality-reduction techniques like the 
discrete Fourier transform \cite{timeseries}.

The method of Wu, et al. \cite{svd-similarity}, conceptually very similar to our 
approach, uses the singular value decomposition (SVD) to find all pairs of close 
vectors with respect to a distance metric (instead of correlation). Similarly, 
Hero and Rajaratnam\cite{hero2011large} apply projections to the thresholded 
correlation problem with the additional objective of determining a reasonable 
threshold value; their method follows a probabilistic approach.  Our method finds 
all of the most highly-correlated vector pairs with respect to a given threshold 
by pruning along an intuitive path of decreasing variance defined by the data and
computed by a truncated SVD.

Consider a real-valued data matrix $A$ consisting of $m$ observations of $n$
column vectors.  Denote the columns of $A$ as $a_j\in\mathcal{R}^m$ for
$j=1,2,\ldots,n$, $A=[a_1, a_2, \ldots, a_n]\in\mathcal{R}^{m\times n}$, and
let $k=\mathrm{rank}(A)$.  Assume that the mean of each column $a_j$ is zero
and that each column has unit norm $\|a_j\| = 1$. Here and below, $\|\cdot\|$
denotes the Euclidean vector norm. Then the Pearson sample correlation matrix
$\mathrm{cor}(A)=A^TA$.  Note that under these assumptions the sample
correlation and covariance matrices are the same up to a constant multiple.
Section \ref{irlba} illustrates how to relax the unit norm and zero mean
assumptions in practice.

Let $A=U\Sigma V^T$ be a singular value decomposition of $A$, where
$U\in\mathcal{R}^{m\times k}$, $V\in\mathcal{R}^{n\times k}$,
$U^TU = V^TV = I$,
and $\Sigma\in\mathcal{R}^{k\times k}$ is a diagonal matrix with
diagonal entries $\sigma_1 \ge \sigma_2 \ge \cdots \ge \sigma_k > 0$.
%%For
%%later convenience, we write $s_{1:p}=[\sigma_1, \sigma_2, \ldots, \sigma_p]^T$
%%to be the vector of the first $p\le k$ singular values along the diagonal of
%%$\Sigma$.
Denote the columns of $V$ as $V=[v_1, v_2, \ldots, v_k]$, where each
$v_j\in\mathcal{R}^n$, $j=1,2,\ldots,k$.  Then $\mathrm{cor}(A) = A^TA =
V\Sigma^2 V^T$ is a symmetric eigenvalue decomposition of the correlation
matrix with nonzero eigenvalues $\sigma_1^2, \sigma_2^2, \ldots, \sigma_k^2$.
The columns of $V$ form an orthonormal basis of the range of the
correlation matrix $A^TA$.

Let $a_i$ and $a_j$, $1\le i,j\le n$ be any two columns of the matrix $A$
(including possibly the case $i=j$). Denote the correlation of these two
vectors by $\mathrm{cor}(a_i,a_j) = a_j^T a_i$.  The following simple lemma
establishes a relationship between correlation and Euclidean distance between
two zero mean, unit norm vectors.

\begin{lemma}
\label{distlemma}
Let $a_i$ and $a_j$, $1\le i,j\le n$ be any two columns of the matrix
$A$. Then
\begin{equation}\label{cordist}
\mathrm{cor}(a_i,a_j) = a_i^Ta_j = 1 - \|a_i - a_j\|^2/2.
\end{equation}
In particular, for a given correlation threshold $t$, $0<t<1$, if
$\|a_i - a_j\|^2 > 2(1-t)$ then
$\mathrm{cor}(a_i, a_j) < t$.
\end{lemma}
%%\begin{proof}
%%\begin{eqnarray*}
%%\|a_i - a_j\|^2 &=& (a_i - a_j)^T(a_i - a_j) \\
%%&=& a_i^T a_i + a_j^T a_j - 2a_i^T a_j \\
%%&=& \|a_i\|^2 + \|a_j\|^2 - 2a_i^T a_j \\
%%&=& 2 - 2a_i^T a_j.
%%\end{eqnarray*}
%%The lemma follows by solving for $a_i^Ta_j$.
%%\end{proof}

Lemma \ref{distlemma} equates the problem of finding pairs of highly correlated
column vectors from the matrix $A$ with a problem of finding pairs of vectors
sufficiently close together. For example, correlation values larger than 0.99
may only be associated with column vectors from the matrix $A$ whose Euclidean
distance is less than $\sqrt{0.02}$.

The following lemma expresses the Euclidean distance between any two columns of
the matrix $A$ as a weighted sum of projected distances along the SVD
basis vectors forming the columns of $V$.
\begin{lemma}
\label{sumlemma}
Let $a_i$ and $a_j$
be any two columns of the matrix
$A$, $1\le i,j\le n$. Then
\begin{equation}\label{vdist}
\|a_i - a_j\|^2 =
\sigma_1^2 (e_i^Tv_{1} - e_j^Tv_{1})^2 + 
\sigma_2^2 (e_i^Tv_{2} - e_j^Tv_{2})^2 + \cdots + 
\sigma_k^2 (e_i^Tv_{k} - e_j^Tv_{k})^2,
\end{equation}
\end{lemma}
where, here and below, $e_{j}\in\mathcal{R}^n$ represents the $j$th unit basis
vector consisting of $1$ in position $j$ and zeros otherwise.
(Thus, $e_j^Tv_{k}$ is the $j$th position in the vector $v_k$--that is, the
$[j,k]$ entry in the matrix $V$.)

%%\begin{proof}
%%\begin{align*}
%%\|a_i - a_j\|^2 &= (a_i - a_j)^T(a_i - a_j) \\
%%&= a_i^Ta_i + a_j^T a_j - 2a_i^Ta_j \\
%%&= \sigma_1^2 v_{1i}^2 + \sigma_2^2 v_{2i}^2 + \cdots + \sigma_k^2 v_{ki}^2 \\
%%&\phantom{=} + \sigma_1^2 v_{1j}^2 + \sigma_2^2 v_{2j}^2 + \cdots + \sigma_k^2 v_{kj}^2 \\
%%&\phantom{=} -2(\sigma_1^2 v_{1i}v_{1j} + \sigma_2^2 v_{2i}v_{2j} + \cdots + \sigma_k^2 v_{ki}v_{kj}) \\
%%&= \sigma_1^2 (v_{1i}^2 + v_{1j}^2 - 2v_{1i}v_{1j}) \\
%%&\phantom{=} + \sigma_2^2 (v_{2i}^2 + v_{2j}^2 - 2v_{2i}v_{2j}) \\
%%&\phantom{=} + \cdots \\
%%&\phantom{=} + \sigma_k^2 (v_{ki}^2 + v_{kj}^2 - 2v_{ki}v_{kj}) \\
%%&= \sigma_1^2 |v_{1i} - v_{1j}|^2
%% + \sigma_2^2 |v_{2i} - v_{2j}|^2
%% + \cdots
%% + \sigma_k^2 |v_{ki} - v_{kj}|^2.
%%\end{align*}
%%\end{proof}

Note that each term in the sum is nonnegative and the weights $\sigma_j^2$,
$j=1,2,\ldots,k$ are nonincreasing and defined by the covariance structure of
the data.  Given a correlation threshold $t$, $0<t<1$, Equations \ref{cordist}
and \ref{vdist} together suggest that pairs of vectors can be excluded from
consideration whenever their projected distance is too large.  For example if
$\sigma_1^2 (e_i^Tv_{1} - e_j^Tv_{1})^2 > 2(1-t)$, then we can conclude from
just a few scalar values that the column vectors $a_i$ and $a_j$ do not meet the
correlation threshold $t$.
In practice, many pairs of vectors may be pruned in this fashion by
inspecting distances in low-dimensional subspaces, substantially reducing the
complexity of computing thresholded correlation matrices.

\section{Efficient pruning using truncated SVD} \label{sect:algo}

Let $t$ be a given correlation threshold, $0<t<1$, and let
$P\in\mathcal{R}^{n\times n}$ be a permutation matrix that orders the entries
of $v_1$ in increasing order.  For example, Figure \ref{fig1} displays the
ordered entries of $Pv_1$ for the small ``EisenYeast'' example (where
$A\in\mathcal{R}^{80 \times 6221}$) from Section \ref{examples}.  The lines in
the plot illustrate the interval associated with the correlation threshold
$t=0.99$ placed at an arbitrary vertical axis location.

\begin{figure}[!ht]
\begin{center}
\includegraphics{draft-001}
\caption{
Example plot of the ordered entries of $v_1$ for the EisenYeast example from
Section \ref{examples}. The red lines illustrate the interval corresponding
threshold $t=0.99$ placed at an arbitrary location on the vertical axis. Points
farther apart than this interval correspond to column vectors that do not meet
the correlation threshold.
\label{fig1}
}
\end{center}
\end{figure}

Let $D^i\in\mathcal{R}^{n-i, n}$, $i=1,2,\ldots,n-1$, be the $i$th order finite
difference matrix with $-1$ along the diagonal, $1$ along the $i$th
super-diagonal, and zeros elsewhere.  Then $D^1P v_1$ consists of the
differences of sequential entries in $P v_1$.  Entries such that $\sigma_1^2 (D^1
P v_1)^2 > 2(1-t)$ correspond to pairs of vectors that do not meet the
correlation threshold, where here and below an exponent applied to a vector
denotes element-wise exponentiation.

Even a 1-d projected interval distance threshold may prune many possible
non-sequential vector pairs with respect to the ordering defined by $P$ as
illustrated by the example in Figure \ref{fig1}.  However, this observation is
unlikely to rule out many sequential pairs of vectors in most problems.  For
instance, the maximum distance between sequential points shown in the example in
Figure \ref{fig1} is less than $\sqrt{2(1-t)}/\sigma_1$, and no pruning of
sequential pairs of vectors with respect to the ordering $P$ occurs.

Including more terms from Equation \ref{vdist} increases our ability to prune
sequential pairs of vectors below the correlation threshold with respect to the
ordering defined by $P$.  Including only one more term finds that 96\% of
sequential vectors fall below the correlation threshold of $0.99$ in the
example shown in Figure \ref{fig1} and are pruned by identifying indices such
that
\[
\sigma_1^2 (D^1 P v_1)^2 +
\sigma_2^2 (D^1 P v_2)^2 > 2(1-t).
\]
Note that we use the permutation $P$ defined by the order of vector $v_1$
throughout.  In general, using $p\le k$ terms to prune pairs of vectors
below the correlation threshold boils down to evaluating the expression
\begin{equation}\label{proj}
(D^1 P V_{1:p} )^2 s_{1:p}^2 > 2(1-t),
\end{equation}
where 
$V_{1:p}$ denotes the first $p$ columns of the
matrix $V$, $s_{1:p}$ the vector of the first $p$ singular values.

Let $\ell$ be the longest run of successive points in $P v_1$ within the
interval $\sqrt{2(1-t)}/\sigma_1$.  The quantity $\ell$ may, for example, be
obtained by rolling the interval over all the points in Figure \ref{fig1} and
counting the maximum number of points that lie within the interval.  With
respect to the ordering $P$, $\ell$ represents the biggest index difference
that pairs of indices corresponding to correlated vectors above the threshold
can exhibit.

Pairs of vectors below the correlation threshold that lie more than one index
difference apart relative to the permutation $P$ can be similarly pruned by
replacing $D^1$ with $D^j$, $j=2,3,\ldots,\ell$ in the above expressions.
Following this process, we can produce a well-pruned candidate set of vector
pairs that contains all pairs that meet a given correlation threshold using
$\ell$ matrix vector products with matrices of order $n \times p$, where $p$ is
chosen in practice such that $p<<k$. Setting $p=10$ for the example shown in
Figure \ref{fig1} with $t=0.99$ prunes all but $149$ possible vector pairs out of a total
$6221(6221 - 1)/2 = 19,347,310$ with $\ell=787$.

These steps are formalized in Algorithm \ref{prune} below. The algorithm
proceeds in two main parts: steps 1--6 prune pairs of vectors down to a small
candidate set that may meet the correlation threshold; step 7
computes the correlation values and applies the threshold across the reduced
set of vector pairs.

\begin{algorithm}\label{prune}
Input: data matrix $A\in \mathcal{R}^{m\times n}$ with columns scaled to have
zero mean and unit norm, correlation threshold $t$,
truncated SVD rank $p<<k$.
\begin{enumerate}
\item Compute the truncated SVD
$AV_{1:p} = U_{1:p}\mathrm{diag}(s_{1:p})$,
where\\
$V_{1:p}=[v_1, v_2, \ldots, v_p]$
and $\mathrm{diag}(s_{1:p})$ denotes the diagonal matrix of the first $p$ singular values.
\item Compute permutation $P$ that orders points in $v_1$.
\item Compute $\ell$, the longest run of successive points in $P v_1$ within the interval $\sqrt{2(1-t)}/\sigma_1$.
\item Compute a set of candidate index pairs $(\tilde{j},\tilde{j}+\tilde{i})$ with respect to the index order defined by the permutation $P$ that possibly meet the correlation threshold
\[
\bigcup_{\tilde{i}=1}^\ell
\left\{
(\tilde{j},\tilde{j}+\tilde{i}) : 
e_{\tilde{j}}^T(D^{\tilde{i}} P V_{1:p} )^2 s_{1:p}^2 \le 2(1-t)
\right\},
\]
where $\tilde{j}=1,2,\ldots,n-\tilde{i}$.
\item If there are judged to be  too many candidate pairs, increase $p$,
compute $AV_{1:p} = U_{1:p}\mathrm{diag}(s_{1:p})$ and go to step 4.
Continue to increase $p$ in this way until the number of candidate
pairs is sufficiently small or stops decreasing (the best this algorithm can do).
\item Recover the non-permuted candidate column vector
indices $(j,i)$ from $j=e_{\tilde{j}}^T Pq$ and $i=e_{\tilde{j}+\tilde{i}}^TPq$ for every
pair $(\tilde{j},\tilde{j} + \tilde{i})$ in step 4, where
$q\in\mathcal{R}^n = [1,2,\ldots,n]^T$.
\item Compute the full correlation coefficient for each column vector pair identified in
the previous step, applying the correlation threshold to this reduced set of values.
\end{enumerate}
\end{algorithm}
One can cheaply estimate whether there are too many candidate pairs for a
given dimension $p$ in step 5 by evaluating step 4 for only sequential pairs of
vectors, $\tilde{i}=1$.

Algorithm \ref{prune} guarantees that no pruned vector pair exceeds the given
correlation threshold--the pruning does not produce false negatives.  The
algorithm typically prunes the vast majority of pairs of vectors below the
correlation threshold with truncated singular value decompositions of only
relatively low rank $p$.  The choice of $p$ represents a balance between work
in computing a truncated SVD and evaluating $\ell$ rank $p$ matrix products in
Step 4 (each increasing in work as $p$ increases), and the work required to
filter the thresholded values in step 7 (decreasing work as $p$ increases).

Most of the floating point operations occur in step 4 of Algorithm \ref{prune}.
Fortunately the $\ell$ evaluations in step 4 are independent of each other and
can easily be computed in parallel. See the reference \proglang{R}
implementation \cite{sup} for an example.

\subsection{General matrices and fast truncated SVD}\label{irlba}

Three practical challenges need to be addressed for Algorithm \ref{prune} to be
used in practice:
\begin{enumerate}
\item The algorithm needs a computationally efficient way to compute
a low-rank truncated SVD of potentially large sparse or dense matrices.
\item The truncated SVD method should be able to cheaply restart to
compute additional singular vectors as needed by step 5 of Algorithm
\ref{prune}.
\item Algorithm \ref{prune} as stated assumes a standardized matrix;
it needs to be extended to efficiently apply to general matrices.
\end{enumerate}
%For Algorithm \ref{prune} to be used in practice, three challenges need to be 
%addressed. First, the algorithm requires an efficient method for computing a 
%relatively low-rank truncated SVD of potentially large matrices required by step 
%1. Second, the truncated SVD method should be able to cheaply restart to compute 
%additional singular vectors as required in step 5 of Algorithm \ref{prune}. 
%Third, while the algorithm assumes a standardized matrix, it should work for 
%general matrices. Fortunately, each of these challenges is met with a single 
%algorithm.

The first two challenges are addressed directly by the augmented implicitly
restarted Lanczos bidiagonalization algorithm (IRLBA) of Baglama and Reichel
\cite{irlba}. It efficiently computes truncated singular value decompositions
of either large dense or sparse matrices and the result from an existing rank
$k$ approximation can be efficiently extended to higher-rank approximations by
restarting.  Reference implementations are available for \proglang{R}
\cite{irlbar}, \proglang{Matlab} \cite{irlbam}, and \proglang{Python}
\cite{irlbap}.

The third challenge is addressed by relaxing the column mean and scale
assumptions without introducing significant additional computational or storage
overhead as follows. Assume $A$ is an $m\times n$ real-valued matrix without
any constant-valued columns. Let $z=[z_1,z_2,\ldots,z_n]\in\mathcal{R}^n$
represent the vector of column means of $A$,
$w=[w_1^2,w_2^2,\ldots,w_n^2]\in\mathcal{R}^n$ be the vector of squared column
norms of $A$, and $W\in\mathcal{R}^{n\times n}$ a diagonal matrix with diagonal
entries $1/\sqrt{w_i^2 - m z_i^2}$.  Let $e\in\mathcal{R}^m$ be a vector of all
ones.  Then $(A - ez^T)W$ is a centered matrix with zero column means and
scaled to have unit column norms, and
\[
\mathrm{cor}(A) = W^T (A-ez^T)^T (A-ez^T) W.
\]
%The IRLBA, based on the Lanczos process, is a method of iterated matrix-vector
%products.
IRLBA does not directly require the matrix $A$, but instead only needs to
compute matrix-vector products with it.  The main idea behind efficient
application of IRLBA to correlation problems replaces matrix-vector products of
the form $Ax$ with $AWx - ez^TWx$, implicitly working with a scaled and
centered matrix without explicit standardization. This comes at the cost of
storing two additional length $n$ vectors and at most computing two additional
vector inner products per matrix-vector product.  The \proglang{R}
implementation \cite{irlbar} of IRLBA includes arguments for centering and
scaling the input matrix.

A na\"ive brute force thresholded correlation computation requires $O(n^2 m)$
flops.  Let $p$ be the selected IRLBA dimension. Then Algorithm \ref{prune}
requires $O(n p \ell)$ flops, where $\ell$ is the longest run of ordered
entries in $v_1$ that meet the 1-d projected distance threshold
$\sqrt{2(1-t)}/\sigma_1$, including the flops required by IRLBA and final
thresholding in step 7.  The main computational savings of $\frac{nm}{lp}$
depends on the rate of decay in the singular values of $A$, the subspace
dimension $p$, and the desired threshold value. We remark that this estimate of
computational savings is available already in step 3 of Algorithm \ref{prune}
after only modest computational effort. Users might use this cheap estimate to,
for example, alter their threshold in some cases.

\section{Extensions}\label{extensions}

A number of practical extensions apply to the main thresholded correlation
idea, most of which are available in the reference R package implementation.
Extensions include finding anti-correlated vectors, finding thresholded
pairwise correlations between column vectors from two different matrices,
application to other distance metrics in place of correlation, and, similarly
to the work of Xiong et al. \cite{prune2}, efficiently finding the top N most
correlated vector pairs for a given constant N.

All of the discussed extensions are simple to implement. This section
outlines key ideas behind the extensions.

\subsection{Spearman correlation and application to other distance metrics}

The idea of evaluating threshold conditions by projection into low dimensional
subspaces applies directly to the Euclidean distance metric, see
\cite{svd-similarity} for a related approach. Lemma \ref{sumlemma} decomposes
Euclidean distance between two vectors in terms of singular basis vectors and
Algorithm \ref{prune} can be applied directly to the problem of finding all
pairs of vectors that meet a thresholded Euclidean distance with small changes
in steps 3, 4, and 7.  By equivalence of norms, thresholded Euclidean distance
can be extended to other distance metrics with increased computational cost.
The reference \proglang{R} implementation \cite{sup} includes functions for
thresholded distance and correlation.

Spearman rank correlation may be found by first transforming the data to ranked
values followed by thresholded correlation using Algorithm \ref{prune} on the
ranked values.

\subsection{Anti-correlation}

Some applications are interested in finding both highly correlated and
anti-correlated vectors--that is, for a given threshold $t > 0$, find column
vectors $a_i$ and $a_j$ such that $|\mbox{cor}(a_i, a_j)| \ge t$.
Note that $-\mbox{cor}(a_i, a_j) = \mbox{cor}(a_i, -a_j)$, and that
similarly to Lemma \ref{sumlemma},
\begin{equation*}
\|a_i + a_j\|^2 =
\sigma_1^2 (e_i^Tv_{1} + e_j^Tv_{1})^2 + 
\sigma_2^2 (e_i^Tv_{2} + e_j^Tv_{2})^2 + \cdots + 
\sigma_k^2 (e_i^Tv_{k} + e_j^Tv_{k})^2,
\end{equation*}
where $e_j, v_j, \sigma_j$ are defined as in Lemma \ref{sumlemma}.
Then steps 2--4 of Algorithm \ref{prune} can be modified to include vector pairs meeting an
anti-correlation threshold as follows:
\begin{enumerate}
\item[2.] Let $\hat{v}_1\in\mathcal{R}^{2n}=[v_1; -v_1]$ (concatenate the vectors). Compute permutation
$P$ that orders points in $\hat{v}_1$.
\item[3.] Compute $\ell$, the longest run of successive points in $P \hat{v}_1$ within the interval $\sqrt{2(1-t)}/\sigma_1$. If $\ell > n-1$, then set $\ell = n-1$.
\item[4.] Compute a set of candidate index pairs $(\tilde{j},\tilde{j}+\tilde{i})$ with respect to the index order defined by the permutation $P$ that possibly meet the correlation threshold
\[
\bigcup_{\tilde{i}=1}^\ell
\left\{
(\tilde{j},\tilde{j}+\tilde{i}) :
\begin{array}{l}
e_{\tilde{j}}^T(D^{\tilde{i}} P V_{1:p} )^2 s_{1:p}^2 \le 2(1-t) \mbox{ or }
\\
e_{\tilde{j}}^T(|D^{\tilde{i}}| P V_{1:p} )^2 s_{1:p}^2 \le 2(1-t)
\end{array}
\right\},
\]
where $\tilde{j}=1,2,\ldots,n-\tilde{i}$, and $|D^{\tilde{i}}|$ indicates
the entry-wise absolute value of the matrix $D^{\tilde{i}}$.
\end{enumerate}
The modified step 3 produces a run length $\ell$ very likely to be larger than
in the default case, and step 4 above requires more arithmetic. Thus, including
anti-correlated vectors in the result is more expensive than simply finding
only highly correlated vectors.

Very similarly, we can compute pairwise thresholded correlation between pairs
of column vectors of two distinct matrices by a truncated SVD of a matrix
formed by coulmn-wise concatenation of the two matrices. The resulting
thresholded pairs are then grouped by original column location to consider only
pairs consisting of column vectors from each original matrix.

\subsection{Top $N$ most correlated pairs}

A procedure that finds the top $N$ most correlated pairs for a fixed integer
$N$ proceeds iteratively starting with a small initial value of $p$, for
instance $p = 0.999$.  Recall that the maximum number of pairs meeting a
threshold is available by step 4 of Algorithm \ref{prune}.  Then:

\begin{enumerate}
\item[a.] Reduce $p$ and repeat steps 1 -- 4 of Algorithm \ref{prune} until
the maximum number of pairs meething the threshold $p$ is greater than $N$.
\item[b.] Complete steps 5 -- 7 of Algorithm \ref{prune}. If the number of
pairs meeting the threshold is less than $N$, reduce $p$ and goto step a.
\end{enumerate}

The procedure then relies on the ability to cheaply restart IRLBA to expand an
existing truncated SVD.

\section{Numerical experiments}\label{examples}

We evaluated the algorithm using public gene expression and DNA methylation
datasets from the \proglang{R} {\tt biclust} \cite{biclust} package and from
the Cancer Genome Atlas \cite{gdac}, referred to as TCGA.  Most tests were
performed on a desktop PC with a single AMD A10-7850K 3.7$\,$GHz Athlon
quad-core CPU and 16$\,$GB 1333$\,$MHz unbuffered RAM running the Ubuntu 15.10
GNU/Linux OS. Tests were performed with 64-bit \proglang{R} version 3.2.3 using
double-precision arithmetic and the OpenBLAS library version 0.2.14-1ubuntu1
(based on Goto's BLAS version 1.13) with {\tt OMP\_NUM\_THREADS=1}. The last
experiment includes a test run on a Linux cluster described in that section.

All examples use the reference \proglang{R} implementation, {\tt tcor()}, which can be
installed from the development GitHub repository using the
{\tt devtools} package with:

\begin{Schunk}
\begin{Sinput}
> devtools::install_github("bwlewis/tcor")
\end{Sinput}
\end{Schunk}

The native \proglang{R} {\tt cor()} function is a general function that can
compute several different types of correlation matrices including Pearson,
Spearman, and Kendall and includes a number of options for dealing with missing
values.  Because of its flexibility, the native {\tt cor()} function does not
always compute Pearson correlation matrices in the most
computationally-efficient way.

Correlation matrix computation can benefit substantially from use of optimized
BLAS Level 3 operations (matrix multiplication), achieving high utilization of
available CPU floating point capacity. The {\tt tcor()} function also benefits
from optimized BLAS routines, but not as much, as it mostly makes use of Level
2 operations and Level 3 operations on sequences of smaller problems.

We compare the {\tt tcor()} function with brute-force Pearson correlation
matrix computation computed in the fastest way we can, directly using matrix
multiplication instead of using the more general {\tt cor()} function.

\subsection{Small gene expression example}

The first example is a small proof of concept that provides the data for Figure
\ref{fig1} and establishes our timing and memory use measurement methodology.
Wall-clock times are reported in seconds and peak memory use in excess of
memory required to store the input data matrix is reported in megabytes.

Anyone can quickly run the example to experiment with the algorithm and
compare its output with brute force results. The example uses the {\tt
EisenYeast} data from the {\tt biclust} \proglang{R} package \cite{biclust}.
The data consist of 80 experiments involving the
Saccharomyces Cerevisiae yeast measuring gene expression data across 6,221
Affymetrix probes, represented as an $80\times 6221$ real-valued matrix.  The
example finds all pairs of columns (gene expression measurements) that meet or exceed a
Pearson correlation value of 0.95. The choice of projection dimension $p$ affects
the efficiency of the algorithm. For $p=2$ the number of candidate vector pairs
generated in step 4 of Algorithm \ref{prune} is over 3 million. When $p$ is increased
to $10$, the number of candidate vector pairs drops to less than 13,000.
The reference \proglang{R} implementation includes options for
tuning $p$ and relies on the restart ability of the IRLBA to pick up where it left off and
inexpensively extend an existing truncated SVD. The following example uses $p=10$.
\begin{small}
\begin{Schunk}
\begin{Sinput}
threshold = 0.95
library(tcor)
library(doMC)     # local parallel computation
library(biclust)  # for the BicatYeast data
data(EisenYeast)  # in gene by sample orientation
A = t(EisenYeast) # to correlate genes to genes
# Register a four-core multicore parallel backend
registerDoMC(4)

# Compute using tcor():
m1  = sum(gc()[, 2])  # Memory in megabytes currently in use
t1  = proc.time()
tx  = tcor(A, t=threshold, p=10)
t1  = (proc.time() - t1)[3] # Time in seconds
m1  = sum(gc()[, 6]) - m1    # Peak excess memory use during the test

# Fast brute force full correlation:
m2  = sum(gc()[, 2])
t2  = proc.time()
mu = colMeans(A)
s  = sqrt(apply(A, 2, crossprod) - nrow(A) * mu ^ 2)
cs = scale(A, center=mu, scale=s)  # scale and center the data
cx = crossprod(cs)                 # full correlation matrix
cx  = cx * upper.tri(cx)    # ignore symmetry and diagonal
pairs = which(cx >= threshold, arr.ind=TRUE)
t2  = (proc.time() - t2)[3]
m2  = sum(gc()[, 6]) - m2
\end{Sinput}
\end{Schunk}
\end{small}
Brute-force computation outperforms {\tt tcor()} in this small first example in
speed but consumes substantially more memory.  Each algorithm identifies the
same set of 125 gene expression pairs that exceed the specified correlation
threshold of 0.95, with comparative timing and memory use shown in the
following table.

% latex table generated in R 3.4.0 by xtable 1.8-2 package
% Sat Aug 12 09:35:30 2017
\begin{table}[ht]
\centering
\begingroup\small
\begin{tabular}{lrr}
  \hline
Method & Wall clock time (s) & Peak excess memory use (MB) \\ 
  \hline
tcor & 3.13 & 59.60 \\ 
  brute force & 1.75 & 1075.10 \\ 
   \hline
\end{tabular}
\endgroup
\caption{Gene expression results, threshold=0.95} 
\label{EisenYeast}
\end{table}

\subsection{TCGA DNA methylation example}

Performance benefits from the {\tt tcor()} function become apparent with more
data.  The following example uses calculated beta values from the Illumina
Human Methylation450 BeadChip of 80 Adenoid Cystic Carcinoma (ACC) tumor
samples obtained from the Broad GDAC Cancer Genome Atlas dashboard \cite{gdac}.
See the \proglang{R} package implementation \cite{sup} for examples of
efficiently converting the downloaded sample methylation data into an R matrix.
We chose these data for their public availability as a general example of
computing thresholded correlation among large numbers of vectors,
rather than for any specific biological meaning.

We removed 91,563 columns with all constant or missing values from the data,
leaving an $80\times 394014$ matrix. The example computes pairs of methylation
beta value column vectors with correlation values of 0.99 or more. We used a
{\tt tcor} projection dimension of 10 and the same performance measurement
methodology used in the first example. Despite its low rank and modest data
size, the problem is too large to compute using a na\"ive brute force
correlation method.  (We remark that the matrix multiplication central to the
brute force implementation can be decomposed into a sequence of smaller steps
to fit within given memory constraints. However, the brute force algorithm
would still need to evaluate all possible vector pair
correlation values.)

Algorithm \ref{prune} identified 913,601 pairs of vectors with correlation
values of 0.99 or more in about three hours on our test PC. A brute force
algorithm would need to evaluate the threshold condition for more than 77
billion vector pairs; the projection of Algorithm \ref{prune} limited the
search space to only about 4 million candidate pairs, from which the 913,601
correlated pairs were found.

We illustrate that steps 4--7 of Algorithm \ref{prune} can be computed in
parallel by comparing the above results run on our reference test PC with
results run on a modest GNU/Linux cluster at Paradigm4, Inc. The cluster
consisted of four computers connected by 10 gigabit ethernet. Each computer
contained two Intel Xeon E5-2650 2$\,$GHz CPUs (16 physical CPU cores per
computer) and 64$\,$GB/1,600$\,$MHz ECC registered RAM running CentOS 6.6 and
64-bit \proglang{R} version 3.2.2 with the reference \proglang{R} BLAS library.

No code change was required to run in parallel on the cluster. The reference \proglang{R}
{\tt tcor()} implementation uses \proglang{R}'s {\tt foreach} \cite{foreach} framework to
run on arbitrary parallel configurations (``back ends''). We simply registered
a {\tt doRedis} \cite{doredis} parallel back end with 16 \proglang{R} workers per computer
(64 total \proglang{R} worker processes) prior to invoking {\tt tcor()}.

The results computed on the Linux cluster are marked ``tcor (64 core cluster)''
below. The cluster example computed the filtering step 7 of Algorithm
\ref{prune} in parallel across the 64 remote \proglang{R} processes. That approach incurs,
in this example, an additional memory overhead of about 16$\,$GB because each
remote \proglang{R} process requires access to a copy of the data matrix $A$. More thrifty
adaptations of the algorithm are possible and could, for example, share the
memory required by $A$ between worker \proglang{R} processes on each computer.

% latex table generated in R 3.4.0 by xtable 1.8-2 package
% Sat Aug 12 09:35:32 2017
\begin{table}[ht]
\centering
\begingroup\small
\begin{tabular}{lrrl}
  \hline
Method & Wall clock time (s) & Memory use (MB) &   \\ 
  \hline
tcor (4 core PC) & 10309.00 & 389.10 &  \\ 
  tcor (64 core cluster) & 482.00 & 994.30 & (local) \\ 
   &  & 42282.40 & (remote) \\ 
   \hline
\end{tabular}
\endgroup
\caption{TCGA ACC Methylation450 (80 rows x 394,014 columns, threshold=0.99). Linux cluster results include master R process memory use (local) plus total peak memory used by all 64 worker R processes (remote).} 
\label{TCGA3}
\end{table}From a set of almost 80 billion pairs of vectors, Algorithm \ref{prune} finds 913,601
pairs with correlation values of 0.99 or more on a modest Linux
cluster in about 8 minutes. The same algorithm works, more slowly, in
the constrained memory setting of a desktop PC without modification.

\section{Conclusion}\label{conclusion}

We present a simple algorithm and reference \proglang{R} implementation based
on a truncated singular value decomposition (SVD) that efficiently computes
thresholded subsets of large correlation  matrices.  We show that the natural
connection between data covariance and the SVD can result in substantial
computational and memory savings, and illustrate these savings with a few
genomics data experiments.  The main computational work in the algorithm is
easily parallelized and deployed across computational clusters.

\bibliography{bibliography}{}
\bibliographystyle{plain}

\end{document}